---
title: "Basic Identification and Visualisation of Outliers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Identification and Visualisation of Outliers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
### Data creation

First, we’re going to load the package and create some outliers in the 'mtcars' dataset. Two continuous variables (one with NA´s), one logical and one factor level. 

```{r setup}
library(outlieR)

set.seed(123)
mtcars["var_1"] <- c(rnorm(29), -100, 100, NA)
set.seed(234)
mtcars["var_2"] <- c(-50, 32, rnorm(30))
mtcars["var_3"] <- c(rep(TRUE, 31), FALSE)
mtcars["var_4"] <- factor(rep(LETTERS[1:5], times = c(10, 5, 15, 1, 1)))

# For better printing.
mtcars <- tibble::as_tibble(mtcars)
```
## Identification of Outliers

The 'identify_outlier()' function is used to identify outliers based on methods and parameters spesificed in 'control_outlier()' function. This control function is optional to use. 


```{r}
outlier_mtcars <-
  identify_outlier(mtcars,
    var_1, var_2, var_3, var_4,
    control = control_outlier(discrete_method = "prop", prop = 0.06, na_action = "keep")
  )
print(outlier_mtcars)
```

To get a summary of the results. 

```{r}
summary(outlier_mtcars)
```


### Filtering

To remove rows in the data with outliers identified in the previous step - call 'filter_outlier()' on the outlier object. 
```{r}
filtred_mtcars <- filter_outlier(outlier_mtcars)

pillar::glimpse(filtred_mtcars)
```

## Visualisation

First continuous 
```{r}
# type  = "histogram" is the default
identify_outlier(mtcars, var_1) |>
  autoplot()

identify_outlier(mtcars, var_1) |>
  autoplot(type = "count")
```

Also multiple 
(not all combinations are supported yet)

```{r}
identify_outlier(mtcars, var_1, var_2) |>
  autoplot(type = "hist")

identify_outlier(mtcars, var_1, var_2, var_3) |>
  autoplot(type = "count")
```

Autoplot returns a ggplot, so its customisable with ggplot2s functions. 

```{r}
identify_outlier(mtcars, var_4) |>
  autoplot(type = "count") + ggplot2::ggtitle("Look E and D are Outliers")
```

